Opportunities for improvement in adult trauma patients admitted to the intensive care unit: A registry-based study

### 2024-06-09 manuscript2.Rmd
9/6

---
title: "Opportunities for improvement in adult trauma patients admitted to the intensive care unit: A registry-based study"
output: "word_document"
author: "Elin Sun Cao"
bibliography: bibliography.bib
csl: vancouver.csl
editor_options: 
  markdown: 
    wrap: 72
---

@@

### Results
Opportunities for improvement was identified in 143 (9,9%) out of 1449 patients. Mechanical ventilation > 7 days (OR 0.35; 95% CI 0.10, 0.89; p=0.05), Revised Trauma Score (OR 1.21; 95% CI 1.08, 1.37; p=0,002), > 7 days in the ICU (OR 1.93; 95% CI 1.23, 3.03; p=0.004) and survival 30 days after injury (OR 3.29; 95% CI 1.60, 7.54; p=0.002) were significantly associated with opportunities for improvement.

@@

# Introduction
Trauma — the clinical entity of external injury and the body’s associated response — is a leading cause of mortality and morbidity worldwide [@60][@6]. Patients with major trauma often require intensive care unit (ICU) admission, where managing these patients can be challenging and can result in complications or death [@64][@69]. Although these complications are often expected, the complexity of ICU care necessitates providers to be highly adaptable, which makes the setting prone to mistakes and errors. Peer review is a cornerstone of trauma quality improvement programs aimed at identifying preventable causes of mortality and morbidity in trauma patients [@70], referred to as opportunities for improvement [@42]. The identification of these help derive corrective action plans to improve the quality of trauma care. Previous research on opportunities for improvement in trauma care has often focused on the initial management, identifying areas such as airway and ventilator management, fluid resuscitation, delays to treatment and judgement errors as common opportunities for improvement [@72][@73][@74].

Assessing preventable deaths among ICU-admitted patients poses a significant challenge due to their medical complexity, making it difficult to establish the standard of care for each individual. Moreover, despite receiving optimal treatment, these patients remain at elevated risk for complications and increased mortality rates [@9]. While previous studies have explored various facets of trauma care, including the ICU-setting [@50], there is a notable gap in research regarding error reporting practices in trauma care - especially for trauma patients admitted to the ICU [@64][@12]. Understanding and addressing this gap is essential for improving patient outcomes and optimizing trauma care protocols. Therefore, this study aims to characterize opportunities for improvement in adult trauma patients admitted to the ICU and assess how patient and process factors are associated with opportunities for improvement in these patients.

@@

The Karolinska University Hospital in Solna is the trauma center designated to receive all severely injured patients in the wider metropolitan area of Stockholm [@17]. The hospital has direct access to radiology, surgery, intensive care, and consultants from relevant specialities [@18]. The care of adult trauma patients is screened for opportunities for improvement using a combination of audit filters and individual review by specialized nurses. Those identified as having a higher likelihood of opportunities for improvement are discussed in a multidisciplinary peer-review held every 6-8 weeks. The presence or absence of opportunities for improvement is decided among the participants of the review and subsequently documented in the trauma care quality database. Before 2017 a non-systematic selection from clinicians involved in the opportunities for improvement process was done by filtering for patients that were thought to have a high risk for opportunities for improvement, but starting in 2017 the care of all patients is screened.

@@

```{r, include=FALSE}
#source("functions/ofi_categories.R")
```

@@

In the adjusted analysis the following factors demonstrated significant association with opportunities for improvement: mechanical ventilation > 7 days `r inline_text(table3b, variable = Intubation, level = "Mechanical ventilation > 7 days")`, Revised Trauma Score `r inline_text(table3b, variable = RTS)`, > 7 days in the ICU `r inline_text(table3b, variable = daysinICU, level = "> 7 days")` and survival 30 days after injury `r inline_text(table3b, variable = Survival, level = "Alive")`.

Sex, age, Injury Severity Score, time to first computed tomography, on-call hours, and ASA preinjury were not significantly associated with opportunities for improvement in the unadjusted nor the adjusted analyses.

@@

Table 4 shows the adjusted associations of opportunities for improvement and the selected patient and process factors among the patients who were alive and dead respectively 30 days after injury. Revised Trauma Score was significantly associated with increased odds of OFI for both patients who were alive `r inline_text(table3aalive, variable = RTS)` and deceased `r inline_text(table3bdead, variable = RTS)` 30 days after injury. More than 7 days in the ICU was significantly associated with increased odds of opportunities for improvement among patients who were deceased 30 days after injury `r inline_text(table3bdead, variable = daysinICU, level = "> 7 days")`.

@@

The key results from this study are that higher Revised Trauma Score, > 7 days length of stay in the ICU, and survival 30 days after injury were significantly associated with higher odds of opportunities for improvement. More than 7 days on mechanical ventilation was associated with decreased odds of opportunities for improvement. Age, sex, Injury Severity Score, time to first computed tomography, arrival during on-call hours, and ASA preinjury were not significantly associated with opportunities for improvement in neither the adjusted nor the unadjusted analyses. In the analysis of patients who were alive and deceased 30 days after injury increased Revised Trauma Score was significantly associated with higher odds of opportunities for improvement in both patient groups. Among the patients who deceased 30 days after injury, more than 7 days in the ICU was significantly associated with higher odds of opportunities for improvement.

@@

We found that mechanical ventilation more than 7 days was significantly associated with decreased odds of opportunities for improvement. Previous research has shown that prolonged mechanical ventilation is associated with increased mortality. This is because mechanical ventilation is associated with severe complications and failure of weaning further increases length of mechanical ventilation as well as ICU length of stay [@65][@66]. Our finding of decreased odds of OFI may be attributed to the fact that, despite the increased rate of complications and mortality associated with prolonged mechanical ventilation, it was the appropriate intervention for these patients.

@@

The findings revealed a significant association between increased odds of opportunities for improvement and higher Revised Trauma Score, ICU length of stay exceeding 7 days, and survival 30 days after injury, as well as a significant association between decreased odds of opportunities for improvement and more than 7 days on mechanical ventilation. The results highlight the importance of prioritizing these groups of patients for closer monitoring and implementing targeted interventions to prevent errors in care. Further research on opportunities for improvement among ICU-admitted patients is needed to expand our understanding of this aspect of trauma care. The primary limitation of this study was its restricted sample size, highlighting the need for larger-scale investigations. Conducting multi-center studies encompassing extensive databases or merging multiple datasets could provide more comprehensive insights into opportunities for improvement and contribute to enhancing Trauma Quality Improvement Programs on a broader scale. However, it is important to acknowledge that multi-center studies may compromise internal validity, reducing their applicability to specific centers.

>>>>

---
title: "Opportunities for improvement in adult trauma patients admitted to the intensive care unit: A registry-based study"
output: "html_document"
author: "Elin Sun Cao"
bibliography: bibliography.bib
csl: vancouver.csl
editor_options: 
  markdown: 
    wrap: 72
---

@@

### Results
Opportunities for improvement was identified in 143 (9,9%) out of 1449 patients.Revised Trauma Score , > 7 days in the ICU  and survival 30 days after injury  were significantly associated with increased opportunities for improvement.

@@

# Introduction
Trauma — the clinical entity of external injury and the body’s associated response — is a leading cause of mortality and morbidity worldwide [@60; @6]. Patients with major trauma often require intensive care unit (ICU) admission, where managing these patients can be challenging and can result in complications or death [@64, @69]. Although these complications are often expected, the complexity of ICU care necessitates providers to be highly adaptable, which makes the setting prone to mistakes and errors. Peer review is a cornerstone of trauma quality improvement programs aimed at identifying preventable causes of mortality and morbidity in trauma patients [@70], referred to as opportunities for improvement [@42]. The identification of these help derive corrective action plans to improve the quality of trauma care. Previous research on opportunities for improvement in trauma care have been compiled in a systematic review, which classified and identified areas for improvement such as inadequate monitoring, team communication, assessment of injury and procedure-related issues [@75].

Assessing preventable deaths among ICU-admitted patients poses a significant challenge due to their medical complexity, making it difficult to establish the standard of care for each individual. Moreover, despite receiving optimal treatment, these patients remain at elevated risk for complications and increased mortality rates [@9]. While previous studies have explored various facets of trauma care, including the ICU-setting [@50], there is a notable gap in research regarding error reporting practices in trauma care - especially for trauma patients admitted to the ICU [@64; @12]. Understanding and addressing this gap is essential for improving patient outcomes and optimizing trauma care protocols. Therefore, this study aims to characterize opportunities for improvement in adult trauma patients admitted to the ICU and assess how patient and process factors are associated with opportunities for improvement in these patients.

@@

The Karolinska University Hospital in Solna is the trauma center designated to receive all severely injured patients in the wider metropolitan area of Stockholm [@17]. The hospital has direct access to radiology, surgery, intensive care, and consultants from relevant specialities [@18]. The care of adult trauma patients is screened for opportunities for improvement using a combination of audit filters and individual review by specialized nurses. Those identified as having a higher likelihood of opportunities for improvement are discussed in a multidisciplinary peer-review held every 6-8 weeks. The presence or absence of opportunities for improvement is decided among the participants of the review and subsequently documented in the trauma care quality database. Prior to 2017 a semi-systematic selection off clinicians involved in the opportunities for improvement process was done by filtering for patients who were thought to have a high risk for opportunities for improvement. However, starting in 2017, the care of all patients is screened using audit filters.

@@

```{r, include=FALSE}
source("functions/ofi_categories.R")
```
```{r, echo=FALSE}
table2
```

@@

In the adjusted analysis the following factors demonstrated significant association with opportunities for improvement: Revised Trauma Score `r inline_text(table3b, variable = RTS)`, > 7 days in the ICU `r inline_text(table3b, variable = daysinICU, level = "> 7 days")` and survival 30 days after injury `r inline_text(table3b, variable = Survival, level = "Alive")`.

Sex, age, mechanical ventilation, Injury Severity Score, time to first computed tomography, on-call hours, and ASA preinjury were not significantly associated with opportunities for improvement in the unadjusted nor the adjusted analyses.

@@

Table 4 shows the adjusted associations of opportunities for improvement and the selected patient and process factors among the patients who were alive and dead respectively 30 days after injury. Revised Trauma Score was significantly associated with increased odds of OFI for both patients who were alive `r inline_text(table3aalive, variable = RTS)` and deceased `r inline_text(table3bdead, variable = RTS)` 30 days after injury.

@@

The key results from this study are that higher Revised Trauma Score, > 7 days length of stay in the ICU, and survival 30 days after injury were significantly associated with higher odds of opportunities for improvement. Age, sex, mechanical ventilation, Injury Severity Score, time to first computed tomography, arrival during on-call hours, and ASA preinjury were not significantly associated with opportunities for improvement in neither the adjusted nor the unadjusted analyses. In the analysis of patients who were alive and deceased 30 days after injury increased Revised Trauma Score was significantly associated with higher odds of opportunities for improvement in both patient groups. Among the patients who deceased 30 days after injury, more than 7 days in the ICU was significantly associated with higher odds of opportunities for improvement.

@@

@@

The findings revealed a significant association between increased odds of opportunities for improvement and higher Revised Trauma Score, ICU length of stay exceeding 7 days, and survival 30 days after injury. The results highlight the importance of prioritizing these groups of patients for closer monitoring and implementing targeted interventions to prevent errors in care. Further research on opportunities for improvement among ICU-admitted patients is needed to expand our understanding of this aspect of trauma care. The primary limitation of this study was its restricted sample size, highlighting the need for larger-scale investigations. Conducting multi-center studies encompassing extensive databases or merging multiple datasets could provide more comprehensive insights into opportunities for improvement and contribute to enhancing Trauma Quality Improvement Programs on a broader scale. However, it is important to acknowledge that multi-center studies may compromise internal validity, reducing their applicability to specific centers.
